Robust Spectroscopic Inference with Imperfect Models

### 2014-11-03 tex/paper.tex
Ready for SA edits

\capstartfalse
\begin{deluxetable}{lr@{ $\pm$ }l|r@{ $\pm$ }l}[!b]
\tablecaption{\label{table:Kurucz}Inferred Parameters for WASP-14}
\tablehead{\colhead{Parameter} & \multicolumn{2}{c}{{\sc CfA/Kurucz}} & \multicolumn{2}{c}{{\sc Phoenix}}}
\startdata
$T_{\rm eff}$ (K) & 6523     & 14                      &  6234   & 17        \\
$\log g$      & 4.29     & 0 (fixed)               &  4.29   & 0 (fixed) \\
$\Z$          & -0.28    & 0.01                    & -0.52  & 0.02     \\
$v \sin i$ ($\kms$)\tablenotemark{a}   & 4.28     & 0.06                    &  4.89   & 0.06      \\
$v_z$ ($\kms$)        & -4.62    & 0.02                    & -4.86   & 0.02      \\
$\log \Omega$\tablenotemark{b} & -12.7180 & 0.0004 & -19.676 & 0.001 \\
$A_V$                          & 0        & 0 (fixed) & 0 & 0 (fixed)
\enddata
\tablenotetext{a}{Small differences in the rotational broadening between the two libraries are
expected, since they assume different micro- and macro-turbulence behavior in the stellar
atmospheres.}
\tablenotetext{b}{The {\sc CfA/Kurucz} spectra are scaled such that the average flux is unity, rather than given in units of flux at the stellar surface, so this value is an unknown multiple of $R^2/d^2$.  The inference on $\Omega$ from the {\sc Phoenix} library is in steradian units.}
\tablecomments{The quoted values represent the ``best-fit", the peak of the marginal posterior 
distributions for each parameter.  The quoted uncertainties correspond to the 68.3\%\ 
($\sim$1\,$\sigma$) confidence intervals.  Expanded lists of the inferred values and uncertainties 
for all nuisance parameters are available electronically.}
\end{deluxetable}
\capstarttrue

@@

Aside from the structure of the covariance matrix, there is also a striking difference between the 
posterior distributions of the stellar parameters inferred using the {\sc Phoenix} and {\sc 
CfA/Kurucz} libraries: see Table~\ref{table:Kurucz} and compare Figures~\ref{fig:Kurucz_posterior} 
and \ref{fig:PHOENIX_posterior}.  The {\sc Phoenix} models find a systematically cooler temperature 
(by $\sim$240\,K) and lower metallicity (by a factor of $\sim$2) than inferred from the {\sc 
CfA/Kurucz} models.  Because our framework explicitly mitigates the potential for posterior bias 
from imperfect matches between data and models, we can safely conclude that these apparent 
discrepancies are intrinsic to the different input physics of the model libraries.\footnote{It is 
worth noting that these differences are not a manifestation of the stringent prior on surface 
gravity.  When the spectra are modeled with $\log g$ unconstrained, the posteriors for both 
libraries shift, but their respective discrepancies actually increase slightly.}  Such a comparison 
serves as a stark reminder that incomplete and/or over-simplified physical prescriptions remain the 
most significant source of systematic uncertainty in estimates of fundamental stellar parameters.

\begin{figure*}[!htb]
  \includegraphics{figs/residuals_Gl51_logg.pdf}
\vspace{-0.3cm}
  \figcaption{The $K$-band SPEX spectrum of Gl~51 ({\it blue}) compared with a {\sc Phoenix} model
({\it red}) generated by drawing parameters from the posterior distribution (for a fixed $\log g = 
5.0$).  ({\it bottom}) The residual spectrum along with contours representing the distributions of
a large number of random draws from the covariance matrix (the shading is representative of the 1,
2, and 3\,$\sigma$ spreads of that distribution of draws), as in Fig.~\ref{fig:Kurucz_residuals}.  
Note how the ``outlier" features are clearly identified and treated by the local covariance kernels.
\label{fig:Gl51_residuals}}
\vspace{0.2cm}
\end{figure*}

@@

By modeling the $K$-band region of the SPEX data with the {\sc Phoenix} library\footnote{The 
differences between the \citet{husser13} {\sc Phoenix} library and the {\sc BT-Settl} modification 
of the {\sc Phoenix} models at these (relatively) high effective temperatures are minor.  Since the 
former is defined over a well-sampled, more regular grid of stellar parameters, we prefer it here 
for computational simplicity.} and a $\delta$-function prior on $\log g$ (at 5.0), we infer 
parameters for Gl~51 that are consistent with \citet{rojas-ayala12}.  
Figure~\ref{fig:Gl51_residuals} displays the spectral modeling results, and 
Figure~\ref{fig:Gl51_posterior} shows the marginal posterior distributions for the relevant stellar 
parameters.  The best-fit parameter values and their associated uncertainties are compiled in 
Table~\ref{table:Gl51}.

We concur with \citet{rojas-ayala12} that the {\sc Phoenix} models manage to match the overall 
spectral morphology of mid-M type stars well, but consistently under-predict the strengths of the 
\ion{Na}{1} and \ion{Ca}{1} resonance lines, even for high metallicities.  \citet{rajpurohit10} 
speculated that this discrepancy may be a consequence of inaccurate atomic data (oscillator 
strengths and/or opacities).  The indexing technique adopted by \citeauthor{rojas-ayala12} presumes 
that these discrepancies do not affect the line {\it ratios}.  Our approach offers an alternative 
that makes no such assumption, instead using local covariance kernels to self-consistently 
determine the ``weights" of discrepant spectral lines in the overall fit.  In that sense, these 
outlier features still bring their full information content to bear on the posterior distributions 
of the stellar parameters without imposing a systematic bias.  Figure~\ref{fig:Gl51_residuals} 
demonstrates that these features have been identified as problematic, resulting in localized 
patches of appropriately inflated covariance.

\capstartfalse
\begin{deluxetable}{lr@{ $\pm$ }l}[h]
\tablecaption{\label{table:Gl51}Inferred Parameters for Gl 51}
\tablehead{\colhead{Parameter} & \multicolumn{2}{c}{{\sc Phoenix}}}
\startdata
$T_{\rm eff}$ (K) & 2998  & 17 \\
$\log g$ & 5.0 & 0 (fixed) \\
$\Z$ & -0.01 & 0.04 \\
$v \sin i$ ($\kms$) & \hili{5.68} & \hili{4.15} \\
$v_z$ ($\kms$) & 12.1 & 1.3 \\
$\log \Omega$ (sr) & -19.628  & 0.003 \\
$A_V$ & 0 & 0 (fixed)
\enddata
\tablecomments{Notation as in Table~\ref{table:Kurucz}.}
\end{deluxetable}
\capstarttrue

@@

Interpolation from a library of synthetic spectra is conceptually straightforward, but in the 
application discussed here there are some subtleties that require detailed consideration.  
Spectral line formation is a non-linear process, where line shapes are determined by the complex 
radiative transfer through a stellar atmosphere.  Any interpolation process aims to bypass this non-linear 
process and predict the behavior of spectral lines based on pre-computed 
library spectra with similar physical parameters.  This appendix describes the 
procedure by which we propagate an empirical approximation of the {\it uncertainty} in that 
prediction--which can be substantial--through the probabilistic framework we have developed.

There are two important factors to consider when interpolating in libraries of model stellar 
spectra.  First, the flux values in some non-negligible fraction of the spectral pixels do not vary 
monotonically with changes in the physical parameters that define the library: this helps inform 
the decision about which functional form of interpolation should be used.  And second, the spectra 
are usually not Nyquist sampled with respect to the physical parameters that define the library.  
This unfortunate fact means that the library is not sampled densely enough in 
parameter-space to reconstruct a synthetic spectrum for arbitrary parameters without some 
information loss.  Given these factors, we choose to develop a spectral ``emulator,'' which, given the information contained in the library of synthesized spectra and a chosen interpolation point, delivers a probability distribution over the range of possible interpolated spectra.

Model library spectra, $f_{\lambda}(\{\vt_{\ast}\}^\textrm{grid})$, are stored as (1-dimensional) arrays of flux 
values, sampled on high resolution wavelength grids (typically $R \gtrsim 100,000$).  In the case 
of interest here, the collections of model parameters $\{\vt_{\ast}\}^\textrm{grid} = [\{T_{\rm eff}, \log g,  \Z 
\}]$ define the dimensions of the library grid; the full library is therefore encapsulated in a 
4-dimensional array.  Both the libraries used here have grid spacings of 0.5 dex in $\log g$ and 
0.5 dex in $\Z$; the {\sc CfA/Kurucz} library steps by 250\,K in $T_{\rm eff}$, but the {\sc 
Phoenix} library has finer coverage in 100\,K increments.

\begin{figure*}[!t]
  \begin{center}
  \includegraphics{figs/pca_reconstruct.pdf}
  \figcaption{ (\emph{top}) The mean spectrum, standard deviation spectrum, and five eigenspectra that form the basis of the \textsc{Phoenix} synthetic library used to model Gl51. (\emph{bottom}) The original synthetic spectrum from the \textsc{Phoenix} library compared with a spectrum reconstructed from a linear combination of the eigenspectra using Eqn~\ref{eqn:reconstruct} with the weights listed in the figure.
\label{fig:pca_reconstruct}}
  \end{center}
\end{figure*}

The first step in designing an emulator is to decompose the spectral library into an appropriate basis for emulation \citep{habib07, heitmann09}. We chose the principal component basis (PCA) to decompose the synthetic library into a set of ``eigenspectra,'' following the techniques of \citet{ivezic13}. Prior to the PCA decomposition, we first identify a subset of $M$ synthetic spectra which will be relevant to the star at hand. We then standardize the synthetic library by first subtracting off the mean spectrum and then ``whiten'' the spectra by dividing by the standard deviation measured in each pixel across the grid. The mean spectrum is 
\begin{equation}
  \xi_\mu = \frac{1}{N} \sum_{i = 1}^N f_\lambda(\{\vt_\ast \}^\textrm{grid}_i)
\end{equation}
and the standard deviation spectrum is
\begin{equation}
  \xi_\sigma = \sqrt{\frac{1}{N} \sum_{i=1}^N \bigl [ f_\lambda(\{\vt_\ast \}^\textrm{grid}_i) - \xi_\mu \bigr]^2 } 
\end{equation}
$\{\vt_\ast \}^\textrm{grid}$ denotes the full collection of stellar parameters sampled in the spectral library, while $\{\vt_\ast \}^\textrm{grid}_i$ denotes a single set of stellar parameters drawn from this collection. All spectra in the library are then standardized by subtracting the mean spectrum and dividing by the standard deviation spectrum
\begin{equation}
  \widetilde{f}_\lambda(\{\vt_\ast \}^\textrm{grid}) = \frac{f_\lambda(\{\vt_\ast \}^\textrm{grid}) - \xi_\mu}{\xi_\sigma}
\end{equation}
Then, the eigenspectra are computed from the standardized grid using PCA analysis (\cite{ivezic13}). We decide to truncate our basis to the first $m$ of the eigenspectra, where $m$ is decided by the minimum number of eigenspectra required to reproduce any spectrum in the grid to better than 2\% accuracy for all pixels; however the mean pixel error is generally much smaller than this ($\lesssim 0.5\%$). We denote a single $N_{\rm pix}$ eigenspectrum as $\xi_k$, with principal component index $k = \{1, 2, \ldots, m\}$. As an example, the eigenspectra basis computed for Gl51 using the \textsc{Phoenix} library is shown in Fig~\ref{fig:pca_reconstruct}, top panel.

Using the PCA basis, we can lossily reconstruct any spectrum from the library with a linear combination of the eigenspectra
\begin{equation}
  f_\lambda(\{\vt_\ast \}^\textrm{grid}_i) \approx \xi_\mu + \xi_\sigma \sum_{k=1}^m w_k(\{\vt_\ast \}^\textrm{grid}_i) \xi_k
  \label{eqn:reconstruct}
\end{equation}
where $w_k$ is the weight of a given eigenspectrum, $k$. This weight is a three dimensional scalar function which depends on the stellar parameters $\vt_\ast$. We can determine the weights at the grid points in the library by taking the dot product of the standardized synthetic spectrum with each of the eigenspectra. The weight for a given eigenspectrum $\xi_k$ is 
\begin{equation}
  w_k(\{\vt_\ast \}^\textrm{grid}_i) = \sum_\lambda \widetilde{f}_\lambda(\{\vt_\ast \}^\textrm{grid}_i) \xi_k
\end{equation}
which is generally a smooth function of stellar parameters (see Figure~\ref{fig:GP_interp}, left panel). To simplify the notation, we can write the weights in a length-$m$ column vector
\begin{equation}
  \mathbf{w}(\{\vt_\ast \}^\textrm{grid}_i) = 
  \begin{bmatrix}
    w_1(\{\vt_\ast \}^\textrm{grid}_i)\\
    w_2(\{\vt_\ast \}^\textrm{grid}_i)\\
    \vdots\\
    w_m(\{\vt_\ast \}^\textrm{grid}_i)
  \end{bmatrix}
\end{equation}
and horizontally concatenate the eigenspectra into a matrix with $N_{\rm pix}$ rows and $m$ columns
\begin{equation}
  \mathbf{\Xi} = 
  \begin{bmatrix}
    \xi_1 & \xi_2 & \cdots & \xi_m \\
  \end{bmatrix}
\end{equation}
Then we can rewrite Equation~\ref{eqn:reconstruct} as
\begin{equation}
  f_\lambda(\{\vt_\ast \}^\textrm{grid}_i) \approx \xi_\mu + \xi_\sigma I_{N_{\rm pix}} \mathbf{\Xi} \mathbf{w}(\{\vt_\ast \}^\textrm{grid}_i)
  \label{eqn:reconstruct2}
\end{equation}
Where $I_{N_{\rm pix}}$ is an $N_{\rm pix}$ by $N_{\rm pix}$ identity matrix. To recapitulate thus far, we have decomposed the spectra in the synthetic library into a principal component basis, allowing us to lossily reconstruct any spectrum in the synthetic library as a weighted linear combination of $m$ eigenspectra. The weight corresponding to each eigenspectrum is a moderately-smooth scalar function of the three stellar parameters. Therefore, to create a spectrum corresponding to a value of stellar parameters which is not in the spectral library, we must interpolate the weights to this new value. In practice, it might be possible to use a traditional interpolation scheme such as spline interpolation to interpolate the weights directly. However, we found that with high-signal to noise spectra (e.g., Gl51 is S/N $> 400$), the error in any given interpolated representation of the spectrum could constitute a significant portion of the uncertainty budget. This, combined with the Nyquist undersampling of the synthetic grid caused behavior where stellar parameters corresponding to a grid point in the synthetic library might be preferentially favored and bias the results, even after empirically accounting for interpolation error by doing ``drop-out'' interpolation tests.

Now, we describe how we design the Gaussian process used for modelling the eigenspectra weights as a function of stellar parameters. In Section~\ref{sec:method}, we used the formalism of Gaussian processes to develop a sophisticated correlated noise model for the data likelihood, which include contributions from a ``global'' covariance kernel to describe the moderate-amplitude correlations present throughout the spectral residuals and several ``local'' covariance kernels to address the highly correlated but-localized spectral residuals originating from pathological errors in the synthetic spectrum. Alternatively, we now use a different Gaussian process in a separate application to serve as an interpolating function for the weights.

We model each eigenspectrum weight independently; therefore there is a separate Gaussian process ($\mathcal{GP}$) for each eigenspectrum. For a single eigenspectrum $k$, we denote the collection of $w_k(\{\vt_\ast \}^\textrm{grid}_i)$ evaluated at all the spectra in the synthetic library as $\mathbf{w}_k^\textrm{grid}$. The Gaussian process treats $\mathbf{w}_k$ as a collection of random variables which are drawn from a single joint multi-variate Gaussian distribution \citep{rasmussen05},
\begin{equation}
  \mathbf{w}_k^\textrm{grid} \sim \mathcal{N} \left ( 0, \mathcal{K}_\textrm{SE} (\{\vt_\ast \}^\textrm{grid}, \{\vt_\ast \}^\textrm{grid})\right )
\end{equation}
The covariance matrix of this multi-variate Gaussian distribution is described by a covariance kernel $\mathcal{K}_\textrm{SE}$. Rather than the \matern\ kernel used in Section~\ref{sec:method}, we elect to use a three dimensional squared-exponential covariance kernel
\begin{multline}
  \mathcal{K}_\textrm{SE}(\vt_{\ast, i}, \vt_{\ast, j} | a_\textrm{SE}, l_\textrm{SE}) = a^2 \exp \left ( 
  -  \frac{(T_{\textrm{eff}_i} - T_{\textrm{eff}_j})^2}{2 l_{T_\textrm{eff}}^2 }  \right ) \times \\
  \exp \left (-\frac{(\log g_i - \log g_j)^2}{2 l_{\log g}^2} \right ) \exp \left ( -\frac{ (\Z_i - \Z_j)^2}{2 l_{\Z}^2} \right )
\end{multline}
This kernel is in contrast to the \matern\ kernel used in Section~\ref{sec:method}, which produces a more jagged Gaussian process reminiscent of the spectral residuals. Then the $M$ by $M$ covariance matrix is 
\begin{equation}
\mathsf{C}^\textrm{grid} = \mathcal{K}_\textrm{SE}( \{\vt_\ast \}^\textrm{grid}, \{\vt_\ast \}^\textrm{grid} | a_\textrm{SE}, l_\textrm{SE})
\end{equation}
which represents the evaluation of the covariance kernel for all pairings of grid point stellar parameters.

>>>>

\capstartfalse
\begin{deluxetable}{lr@{ $\pm$ }l|r@{ $\pm$ }l}[!b]
\tablecaption{\label{table:Kurucz}Inferred Parameters for WASP-14}
\tablehead{\colhead{Parameter} & \multicolumn{2}{c}{{\sc CfA/Kurucz}} & \multicolumn{2}{c}{{\sc Phoenix}}}
\startdata
$T_{\rm eff}$ (K) & 6424     & 19                      &  6354   & 39        \\
$\log g$      & 4.29     & 0 (fixed)               &  4.29   & 0 (fixed) \\
$\Z$          & -0.26    & 0.01                    & -0.30  & 0.02     \\
$v \sin i$ ($\kms$)\tablenotemark{a}   & 4.47 & 0.06       &  5.09   & 0.06      \\
$v_z$ ($\kms$)        & -4.60    & 0.02                    & -4.86   & 0.02      \\
$\log \Omega$\tablenotemark{b} & -12.7180 & 0.0004 & -12.717 & 0.0007 \\
$A_V$                          & 0        & 0 (fixed) & 0 & 0 (fixed)
\enddata
\tablenotetext{a}{Small differences in the rotational broadening between the two libraries are
expected, since they assume different micro- and macro-turbulence behavior in the stellar
atmospheres.}
\tablenotetext{b}{For the spectral emulator (Appendix~\ref{sec:Appendix}) the spectra are scaled such that the average flux is unity, rather than given in units of flux at the stellar surface, so this value is an unknown multiple of $R^2/d^2$.}
\tablecomments{The quoted values represent the ``best-fit", the peak of the marginal posterior 
distributions for each parameter.  The quoted uncertainties correspond to the 68.3\%\ 
($\sim$1\,$\sigma$) confidence intervals.  Expanded lists of the inferred values and uncertainties 
for all nuisance parameters are available electronically.}
\end{deluxetable}
\capstarttrue

@@

Aside from the structure of the covariance matrix, there is also a moderate difference between the 
posterior distributions of the stellar parameters inferred using the {\sc Phoenix} and {\sc 
CfA/Kurucz} libraries: see Table~\ref{table:Kurucz} and compare Figures~\ref{fig:Kurucz_posterior} 
and \ref{fig:PHOENIX_posterior}.  The {\sc Phoenix} models find a systematically cooler temperature 
(by $\sim$75\,K) and lower metallicity than inferred from the {\sc 
CfA/Kurucz} models.  Because our framework explicitly mitigates the potential for posterior bias 
from imperfect matches between data and models, we can safely conclude that these apparent 
discrepancies are intrinsic to the different input physics of the model libraries.\footnote{It is 
worth noting that these differences are not a manifestation of the stringent prior on surface 
gravity.  When the spectra are modeled with $\log g$ unconstrained, the posteriors for both 
libraries shift, but their respective discrepancies actually increase slightly.}  Such a comparison 
serves as a stark reminder that incomplete and/or over-simplified physical prescriptions remain the 
most significant source of systematic uncertainty in estimates of fundamental stellar parameters.

\begin{figure*}[!htb]
  \includegraphics{figs/residuals_Gl51_logg.pdf}
\vspace{-0.3cm}
  \figcaption{The $K$-band SPEX spectrum of Gl~51 ({\it blue}) compared with a
  {\sc Phoenix} model ({\it red}) generated by drawing parameters from the
  posterior distribution. ({\it bottom}) The residual spectrum along with
  contours representing the distributions of a large number of random draws
  from the covariance matrix (the shading is representative of the 1,
2, and 3\,$\sigma$ spreads of that distribution of draws), as in Fig.~\ref{fig:Kurucz_residuals}.  
Note how the ``outlier" features are clearly identified and treated by the local covariance kernels.
\label{fig:Gl51_residuals}}
\vspace{0.2cm}
\end{figure*}

@@

By modeling the $K$-band region of the SPEX data with the {\sc Phoenix} library\footnote{The 
differences between the \citet{husser13} {\sc Phoenix} library and the {\sc BT-Settl} modification 
of the {\sc Phoenix} models at these (relatively) high effective temperatures are minor.  Since the 
former is defined over a well-sampled, more regular grid of stellar parameters, we prefer it here 
for computational simplicity.} and a Gaussian prior on $\log g$ (at $5.0 \pm 0.05$ dex), we infer 
parameters for Gl~51 that are consistent with \citet{rojas-ayala12}.  
Figure~\ref{fig:Gl51_residuals} displays the spectral modeling results, and 
Figure~\ref{fig:Gl51_posterior} shows the marginal posterior distributions for the relevant stellar 
parameters.  The best-fit parameter values and their associated uncertainties are compiled in 
Table~\ref{table:Gl51}.

We concur with \citet{rojas-ayala12} that the {\sc Phoenix} models manage to match the overall 
spectral morphology of mid-M type stars well, but consistently under-predict the strengths of the 
\ion{Na}{1} and \ion{Ca}{1} resonance lines, even for high metallicities.  \citet{rajpurohit10} 
speculated that this discrepancy may be a consequence of inaccurate atomic data (oscillator 
strengths and/or opacities).  Our approach uses local covariance kernels to self-consistently 
determine the ``weights" of discrepant spectral lines in the overall fit.  In that sense, these 
outlier features still bring their full information content to bear on the posterior distributions 
of the stellar parameters without imposing a systematic bias.  Figure~\ref{fig:Gl51_residuals} 
demonstrates that these features have been identified as problematic, resulting in localized 
patches of appropriately inflated covariance.

\capstartfalse
\begin{deluxetable}{lr@{ $\pm$ }l}[h]
\tablecaption{\label{table:Gl51}Inferred Parameters for Gl 51}
\tablehead{\colhead{Parameter} & \multicolumn{2}{c}{{\sc Phoenix}}}
\startdata
$T_{\rm eff}$ (K) & 3071  & 28 \\
$\log g$ & 5.19 & 0.04 \\
$\Z$ & 0.14 & 0.07 \\
$v \sin i$ ($\kms$) & \multicolumn{2}{c}{$\leq 15.0$} \\
$v_z$ ($\kms$) & 11.6 & 1.3 \\
$\log \Omega$ (sr) & -13.481  & 0.001 \\
$A_V$ & 0 & 0 (fixed)
\enddata
\tablecomments{Notation as in Table~\ref{table:Kurucz}.}
\end{deluxetable}
\capstarttrue

@@

Interpolation from a library of synthetic spectra is a conceptually straightforward process, but in the 
application discussed here there are some subtleties that require detailed consideration.  
Spectral line formation is a non-linear process, where line shapes are determined by complex 
radiative transfer through a stellar atmosphere.  Any interpolation process aims to bypass this non-linear 
process and predict the behavior of spectral lines at intermediate points based on pre-computed 
library spectra with similar physical parameters.  This appendix describes the 
procedure by which we propagate an empirical approximation of the {\it uncertainty} in that 
prediction--which can be substantial--through the probabilistic framework we have developed.

There are two important factors to consider when interpolating libraries of model stellar 
spectra.  First, the flux values in some non-negligible fraction of the spectral pixels do not vary 
monotonically with changes in the physical parameters that define the library: this helps inform 
the decision about which functional form of interpolation should be used.  And second, the spectra 
are usually not Nyquist sampled with respect to the physical parameters that define the library.  
This unfortunate fact means that the library is not sampled densely enough in 
parameter-space to reconstruct a synthetic spectrum for arbitrary parameters without some 
information loss.  Given these factors, we choose to develop a spectral ``emulator,'' which, given the information contained in the library of synthesized spectra and a chosen interpolation point, delivers a probability distribution over the range of possible interpolated spectra.

Model library spectra are stored as (1-dimensional) arrays of flux values,
sampled on high resolution wavelength grids (typically $R \gtrsim 100,000$).
In the case of interest here, the collections of model parameters
$\{\vt_{\ast}\}^\textrm{grid} = [\{T_{\rm eff}, \log g,  \Z \}]$ define the
dimensions of the library grid; the full library
$f_{\lambda}(\{\vt_{\ast}\}^\textrm{grid})$ is therefore encapsulated in a
4-dimensional array.  Both the libraries used here have grid spacings of 0.5
dex in $\log g$ and 0.5 dex in $\Z$; the {\sc CfA/Kurucz} library steps by
250\,K in $T_{\rm eff}$, but the {\sc Phoenix} library has finer coverage in
100\,K increments.

\begin{figure*}[!t]
  \begin{center}
  \includegraphics{figs/pca_reconstruct.pdf}
  \figcaption{ (\emph{top}) The mean spectrum, standard deviation spectrum, and five eigenspectra that form the basis of the \textsc{Phoenix} synthetic library used to model Gl51, generated using a subset of the \textsc{Phoenix} grid parameters space most relevant for M dwarfs. (\emph{bottom}) The original synthetic spectrum from the \textsc{Phoenix} library compared with a spectrum reconstructed from a linear combination of the eigenspectra using Eqn~\ref{eqn:reconstruct} with the weights listed in the figure.
\label{fig:pca_reconstruct}}
  \end{center}
\end{figure*}

The first step in designing an emulator is to decompose the spectral library into an appropriate basis for emulation \citep{habib07, heitmann09}. We chose the principal component basis (PCA) to decompose the synthetic library into a set of ``eigenspectra,'' following the techniques of \citet{ivezic13}. Prior to the PCA decomposition, we first isolate a subset of the library (containing $M$ spectra) with parameter values that will be most relevant to the star at hand. We then standardize these spectra by first subtracting off the mean spectrum and then ``whiten'' the spectra by dividing by the standard deviation measured in each pixel across the grid. The mean spectrum is 
\begin{equation}
  \xi_\mu = \frac{1}{M} \sum_{i = 1}^M f_\lambda(\{\vt_\ast \}^\textrm{grid}_i)
\end{equation}
and the standard deviation spectrum is
\begin{equation}
  \xi_\sigma = \sqrt{\frac{1}{M} \sum_{i=1}^M \bigl [ f_\lambda(\{\vt_\ast \}^\textrm{grid}_i) - \xi_\mu \bigr]^2 } 
\end{equation}
where $\{\vt_\ast \}^\textrm{grid}$ denotes the full collection of the $M$ pairings of stellar parameters sampled in the spectral library and $\{\vt_\ast \}^\textrm{grid}_i$ denotes a single set of stellar parameters drawn from this collection. All spectra in the library are then standardized by subtracting the mean spectrum and dividing by the standard deviation spectrum
\begin{equation}
  \hat{f}_\lambda(\{\vt_\ast \}^\textrm{grid}) = \frac{f_\lambda(\{\vt_\ast \}^\textrm{grid}) - \xi_\mu}{\xi_\sigma}
\end{equation}
Then, the eigenspectra are computed from the standardized grid using PCA analysis (\cite{ivezic13}). We decide to truncate our basis to the first $m$ eigenspectra, where $m$ is decided by the minimum number of eigenspectra required to reproduce any spectrum in the grid to better than 2\% accuracy for all pixels; the mean pixel error is generally much smaller than this ($\lesssim 0.5\%$). We denote a single length $N_{\rm pix}$ eigenspectrum as $\xi_k$, with principal component index $k = \{1, 2, \ldots, m\}$. For example, the eigenspectra basis computed for Gl51 using the \textsc{Phoenix} library is shown in Fig~\ref{fig:pca_reconstruct}, top panel.

\begin{figure*}[!t]
\begin{center}
  \includegraphics{figs/GP_left.pdf}
  \includegraphics{figs/GP_right.pdf}
  \figcaption{The Gaussian process modelling of the principal component weights for the Gl51 \textsc{Phoenix} spectral library. (\emph{left}) the blue dots mark the weights $\mathbf{w}_5^\textrm{grid}$ of the 5th eigenspectrum $\xi_5$ computed at a one-dimensional slice of the spectral library for grid points with $\log g = 5.0$, $\Z = 0.0$ and various values of $T_\textrm{eff}$. In reality, the weights are a three-dimensional function of $\vt_\ast$. The thin blue lines show 50 random draws of possible functional forms described by the Gaussian process, constrained to pass through the grid weights. (\emph{inset}) a zoomed portion showing the scatter in the possible functional forms. The black vertical line represents a slice through the scatter of the predicted weight value at $\vt_\ast = [T_\textrm{eff} = 3150\,\textrm{K}, \log g = 5.0\,\textrm{dex}, \Z = 0.0\, \textrm{dex}]$. (\emph{right}) The posterior probability of $w_k$ at this value of $\vt_\ast$ is completely described by Eqn~\ref{eqn:conditional_gp}, allowing us to analytically marginalize over all probable values of the weights, and thus marginalize over all probable spectral interpolations.
\label{fig:GP_interp}}
\end{center}
\end{figure*}

Using the principal component basis, we can lossily reconstruct any spectrum from the library with a linear combination of the eigenspectra
\begin{equation}
  f_\lambda(\{\vt_\ast \}^\textrm{grid}_i) \approx \xi_\mu + \xi_\sigma \sum_{k=1}^m w_k(\{\vt_\ast \}^\textrm{grid}_i) \xi_k
  \label{eqn:reconstruct}
\end{equation}
where $w_k$ is the weight of a given eigenspectrum, $k$. This weight is a three dimensional scalar function which depends on the stellar parameters $\vt_\ast$. We can determine the weight at any grid point in the library by taking the dot product of the standardized synthetic spectrum with the eigenspectrum  
\begin{equation}
  w_k(\{\vt_\ast \}^\textrm{grid}_i) = \sum_\lambda \hat{f}_\lambda(\{\vt_\ast \}^\textrm{grid}_i) \xi_k
\end{equation}
and is generally a smooth function of stellar parameters (see Figure~\ref{fig:GP_interp}, left panel). To simplify notation, we can write the collection of eigenspectra weights in a length-$m$ column vector
\begin{equation}
  \mathbf{w}(\vt_\ast) = 
  \begin{bmatrix}
    w_1(\vt_\ast )\\
    w_2(\vt_\ast )\\
    \vdots\\
    w_m(\vt_\ast )
  \end{bmatrix}
\end{equation}
and horizontally concatenate the eigenspectra into a matrix with $N_{\rm pix}$ rows and $m$ columns
\begin{equation}
  \mathbf{\Xi} = 
  \begin{bmatrix}
    \xi_1 & \xi_2 & \cdots & \xi_m \\
  \end{bmatrix} .
\end{equation}
Then, we can rewrite Equation~\ref{eqn:reconstruct} as
\begin{equation}
  f_\lambda(\{\vt_\ast \}^\textrm{grid}_i) \approx \xi_\mu + \xi_\sigma I_{N_{\rm pix}} \mathbf{\Xi} \mathbf{w}(\{\vt_\ast \}^\textrm{grid}_i)
  \label{eqn:reconstruct2}
\end{equation}
where $I_{N_{\rm pix}}$ is an ($N_{\rm pix}, N_{\rm pix}$) identity matrix.

To recapitulate, we have decomposed the spectra in the synthetic library into a principal component basis, allowing us to lossily reconstruct any spectrum in the synthetic library as a weighted linear combination of $m$ eigenspectra. The weight corresponding to each eigenspectrum is a moderately-smooth scalar function of the three stellar parameters. Therefore, to create a spectrum corresponding to a value of stellar parameters which is not in the spectral library, we must interpolate the weights to this new value. In practice, it may be possible to use a traditional interpolation scheme such as spline interpolation to interpolate the weights directly. However, we found that with high-signal to noise spectra (e.g., Gl51 is S/N $> 400$), the error in any given interpolated representation of the spectrum can constitute a significant portion of the uncertainty budget. This, combined with the Nyquist undersampling of the synthetic grid can cause artificial ``peaking'' of the posterior near grid points in the synthetic library, because the interpolated spectrum is not as good as the raw spectrum at the grid point. Even explicitly accounting for interpolation error by doing ``drop-out'' interpolation tests and empricially propagating it forward did not relieve the peaking problem. We address this problem by using a Gaussian process to encapsulate the range of possible interpolated spectra.

Now, we describe the Gaussian process used for modelling the eigenspectra weights as a function of stellar parameters. In Section~\ref{sec:method}, we used the formalism of Gaussian processes to develop a sophisticated correlated noise model for the data likelihood, which include contributions from a ``global'' covariance kernel to describe the moderate-amplitude correlations present throughout the spectral residuals and several ``local'' covariance kernels to address the highly correlated but-localized spectral residuals originating from pathological errors in the synthetic spectrum.  We now use a separate Gaussian process to serve as an interpolating function for the weights. Because each eigenspectrum weight is modelled independently, there is a new Gaussian process ($\mathcal{GP}$) for each eigenspectrum. For a single eigenspectrum $k$, we denote the collection of $w_k(\{\vt_\ast \}^\textrm{grid}_i)$ evaluated at all the spectra in the synthetic library as $\mathbf{w}_k^\textrm{grid}$. The Gaussian process treats $\mathbf{w}_k$ as a collection of random variables which are drawn from a single joint multi-variate Gaussian distribution \citep{rasmussen05},
\begin{equation}
  \mathbf{w}_k^\textrm{grid} \sim \mathcal{N} \left ( \mathbf{0}, \mathbf{\Sigma}_k^\textrm{grid} \right )
\end{equation}
The covariance matrix of this multi-variate Gaussian distribution is described by a covariance kernel $\mathcal{K}_\textrm{SE}$. Rather than the \matern\ kernel used in Section~\ref{sec:method}, we elect to use a three dimensional squared-exponential covariance kernel
\begin{multline}
  \mathcal{K}_\textrm{SE}(\vt_{\ast, i}, \vt_{\ast, j} | a_\textrm{SE}, l_{T_\textrm{eff}},  l_{\log g}, l_{\Z}) = \\
  a^2 \exp \left ( - \frac{(T_{\textrm{eff}_i} - T_{\textrm{eff}_j})^2}{2 l_{T_\textrm{eff}}^2 }  \right )
  \exp \left (-\frac{(\log g_i - \log g_j)^2}{2 l_{\log g}^2} \right ) \times \\
  \exp \left ( -\frac{ (\Z_i - \Z_j)^2}{2 l_{\Z}^2} \right )
\end{multline}
This kernel is in contrast to the \matern\ kernel used in Section~\ref{sec:method}, which produces a more jagged Gaussian process reminiscent of the spectral residuals. The squared exponential kernel produces a smooth functional form which closely matches the behavior of the eigenspectra weights (Fig~\ref{fig:GP_interp}, left panel). Then the ($M,M$) covariance matrix is 
\begin{equation}
\mathbf{\Sigma}_k^\textrm{grid} = \mathcal{K}_\textrm{SE}( \{\vt_\ast \}^\textrm{grid}, \{\vt_\ast \}^\textrm{grid})
\end{equation}
which represents the evaluation of the covariance kernel for all pairings of grid point stellar parameters, and the conditional dependence on the kernel hyperparameters is implied.
